---
title: "Homework 5"
author: "Akshaya Ravikumar"
date: "02/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/UMBio201/Lab5/")
```

# Load packages
```{r Load packages, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(tidyverse)
library(readxl)
library(broom)
library(cowplot)
set.seed(7)
```

Save any data frames generated to curated_data directory, save all plots to figures directory. Based on assumptions, clearly state which statistical test will be used and why. Include any statistics (p- or R-values) as comments after the tests.

# Question 1

Use the SCFA wide dataframe from Lab 5. Check assumptions, conduct statistical tests, make a plot, to answer the question, "Which supplement causes the greatest increase in acetate concentrations?"
```{r}
# import and filter data set (q1_df)
q1_df <- read_delim(file = "raw_data/scfa_wide.txt", 
                        delim = "\t", escape_double = FALSE, 
                        trim_ws = TRUE, na=c("NA"),
                        col_names = TRUE,
                        col_types = cols(participant_id = col_character(),
                                         semester = col_character(),
                                         supplement_consumed = col_character(),
                                         quantity_compliant = col_character(),
                                         frequency = col_character(),
                                         acetate_delta = col_double(),
                                         butyrate_delta = col_double(),
                                         propionate_delta = col_double())) %>%
  filter(semester != "Winter2015",
         quantity_compliant != "no")
```

```{r}
# check assumptions
bartlett.test(acetate_delta ~ supplement_consumed, data = q1_df)

```

```{r}
# conduct statistical test
kruskal.test(supplement_consumed ~ acetate_delta, q1_df)
```

```{r}
# make and save plot (q1_plot)
q1_plot <- q1_df %>%
    filter(semester != "Winter2015",
         quantity_compliant != "no") %>%
  ggplot(aes(x = supplement_consumed, 
             y = acetate_delta, 
             color = supplement_consumed)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_violin() + 
  geom_jitter() + 
  xlab(NULL) + 
  ylab("Acetate Concentration") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

q1_plot
save_plot(q1_plot, 
          nrow = 2 , ncol = 2, 
          filename = "figures/Q1_aceplot.pdf")


```

What are your conclusions regarding the research question? 

Since the p-value for the Kruskal Wallis test is very high, we may suggest that there is no significant difference among the different types of starch in relation to acetate concentrations.


# Question 2

Use the SCFA wide dataframe from Lab 5. Check assumptions, conduct statistical tests, make a plot, to answer the question, "Which supplement causes the greatest increase in propionate concentrations?"
```{r}
# import and filter data set (q2_df)

q2_df <- read_delim(file = "raw_data/scfa_wide.txt", 
                        delim = "\t", escape_double = FALSE, 
                        trim_ws = TRUE, na=c("NA"),
                        col_names = TRUE,
                        col_types = cols(participant_id = col_character(),
                                         semester = col_character(),
                                         supplement_consumed = col_character(),
                                         quantity_compliant = col_character(),
                                         frequency = col_character(),
                                         acetate_delta = col_double(),
                                         butyrate_delta = col_double(),
                                         propionate_delta = col_double())) %>%
  filter(semester != "Winter2015",
         quantity_compliant != "no")
```

```{r}
# check assumptions
#Equal Variances -- MET
bartlett.test(propionate_delta ~ supplement_consumed, data = q2_df)

# check sample sizes
q2_df %>%
 group_by(supplement_consumed) %>%
  summarise(sample_size = n())

# check normality of each group 
s1 <- q2_df %>%
  filter(supplement_consumed == "Accessible") 
shapiro.test(s1$propionate_delta) #p-value = 0.2301

s2 <- q2_df %>%
  filter(supplement_consumed == "Arabino") 
shapiro.test(s2$propionate_delta) #p-value = 0.9738

s3 <- q2_df %>%
  filter(supplement_consumed == "BRMPS+Accessible")
shapiro.test(s3$propionate_delta) #p-value = 0.09113

s4 <- q2_df %>%
  filter(supplement_consumed == "HiMaize") 
shapiro.test(s4$propionate_delta) #p-value = 0.02262

s5 <- q2_df %>%
  filter(supplement_consumed == "HiMaize+BRMPS") 
shapiro.test(s5$propionate_delta) #p-value = 0.0156

s6 <- q2_df %>%
  filter(supplement_consumed == "Inulin") 
shapiro.test(s6$propionate_delta) #p-value = 0.009586

s7 <- q2_df %>%
  filter(supplement_consumed == "LOODAT")
shapiro.test(s7$propionate_delta) #p-value = 0.8912

s8 <- q2_df %>%
  filter(supplement_consumed == "none")
shapiro.test(s8$propionate_delta) #p-value = 0.009973

s9 <- q2_df %>%
  filter(supplement_consumed == "Psyllium")
shapiro.test(s9$propionate_delta) #p-value = 0.5866

s10 <- q2_df %>%
  filter(supplement_consumed == "BRMPS")
shapiro.test(s10$propionate_delta) #p-value =  7.829e-05

s11 <- q2_df %>%
  filter(supplement_consumed == "transition_HiMaize")
shapiro.test(s11$propionate_delta) #p-value =  0.7647

s12 <- q2_df %>%
  filter(supplement_consumed == "Psyllium+BRMPS")
shapiro.test(s12$propionate_delta) #p-value =  0.03922
```

```{r}
# conduct statistical test
kruskal.test(supplement_consumed ~ propionate_delta, q2_df)
```

```{r}
# make and save plot (q2_plot)
q2_plot <- q2_df %>%
    filter(semester != "Winter2015",
         quantity_compliant != "no") %>%
  ggplot(aes(x = supplement_consumed, 
             y = propionate_delta, 
             color = supplement_consumed)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_violin() + 
  geom_jitter() + 
  xlab(NULL) + 
  ylab("Propionate Concentration") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

q2_plot

save_plot(q2_plot, 
          nrow = 2 , ncol = 2, 
          filename = "figures/Q2_propplot.pdf")

```

What are your conclusions regarding the reseach question?

Due to the high p-value in the Kruskal-Wallis test, we may suggest that there is no significant difference among the different types of starch in relation to propionate concentrations.



# Question 3

Use the SCFA wide dataframe from Lab 5. Subset and filter the data frame, check assumptions, and choose the appropriate statistical tests, and create a plot to answer the following research question: "Does the proportion of acetate reponses differ between individuals who consumed BRMPS potato (BRMPS) or LOODAT potato supplements?"

```{r}
# import and filter data set (q3_df)
q3_df <- read_delim(file = "raw_data/scfa_wide.txt", 
                        delim = "\t", escape_double = FALSE, 
                        trim_ws = TRUE, na=c("NA"),
                        col_names = TRUE,
                        col_types = cols(participant_id = col_character(),
                                         semester = col_character(),
                                         supplement_consumed = col_character(),
                                         quantity_compliant = col_character(),
                                         frequency = col_character(),
                                         acetate_delta = col_double(),
                                         butyrate_delta = col_double(),
                                         propionate_delta = col_double())) %>%
  filter(semester != "Winter2015",
         quantity_compliant != "no",
         supplement_consumed == "LOODAT" | supplement_consumed == "BRMPS") %>%
  select(-butyrate_delta, -propionate_delta) %>% 
  add_column(ace_status = "-") %>%
  drop_na()
```

```{r}
# check assumptions
q3_df %>%
 group_by(supplement_consumed) %>%
  summarise(sample_size = n())

s13 <- q3_df %>%
  filter(supplement_consumed == "BRMPS")
shapiro.test(s13$acetate_delta)
#Low p-value, but large sample size and thus CLT is met.

s14 <- q3_df %>%
  filter(supplement_consumed == "LOODAT")
shapiro.test(s14$acetate_delta)

q3_df$ace_status[q3_df$acetate_delta > 0] <- "positive"
q3_df$ace_status[q3_df$acetate_delta < 0] <- "negative"
```

```{r}
# conduct statistical test
ace_tab <-  with(q3_df, table(supplement_consumed, ace_status))

q3_df %>%
  group_by(supplement_consumed, ace_status) %>%
  summarise(Counts = n())

prop.test(ace_tab, 
          alternative = "two.sided",
          correct = FALSE)
```
| | Increased acetate | Total participants | Percent increase |
|:-----:|:-----:|:-----:|:-----:|
| BRMPS | 75 | 128 | 58.6% |
| LOODAT | 9 | 16 | 56.25% | 

```{r}
# make and save plot (q3_plot)
```

What are your conclusions regarding the reseach question?

Since the p-value for the proportion test was very high, we may suggest that there is no difference in the proportion of acetate reponses between individuals who consumed BRMPS potato (BRMPS) or LOODAT potato supplements.


# Question 4

Use the SCFA wide dataframe from Lab 5. Subset and filter the data frame, check assumptions, and choose the appropriate statistical tests, and create a plot to answer the following research question: "Does the proportion of propionate reponses differ between individuals who consumed BRMPS potato (BRMPS) or LOODAT potato supplements?"

```{r}
# import and filter data set (q4_df)
q4_df <- read_delim(file = "raw_data/scfa_wide.txt", 
                        delim = "\t", escape_double = FALSE, 
                        trim_ws = TRUE, na=c("NA"),
                        col_names = TRUE,
                        col_types = cols(participant_id = col_character(),
                                         semester = col_character(),
                                         supplement_consumed = col_character(),
                                         quantity_compliant = col_character(),
                                         frequency = col_character(),
                                         acetate_delta = col_double(),
                                         butyrate_delta = col_double(),
                                         propionate_delta = col_double())) %>%
  filter(semester != "Winter2015",
         quantity_compliant != "no",
         supplement_consumed == "LOODAT" | supplement_consumed == "BRMPS") %>%
  select(-butyrate_delta, -acetate_delta) %>% 
  add_column(prop_status = "-") %>%
  drop_na()
```

```{r}
# check assumptions
q4_df %>%
 group_by(supplement_consumed) %>%
  summarise(sample_size = n())

s15 <- q4_df %>%
  filter(supplement_consumed == "BRMPS")
shapiro.test(s15$propionate_delta)
#Low p-value, but large sample size and thus CLT is met.

s16 <- q4_df %>%
  filter(supplement_consumed == "LOODAT")
shapiro.test(s16$propionate_delta)

q4_df$prop_status[q4_df$propionate_delta > 0] <- "positive"
q4_df$prop_status[q4_df$propionate_delta < 0] <- "negative"
```

```{r}
# conduct statistical test
prop_tab <-  with(q4_df, table(supplement_consumed, prop_status))

q4_df %>%
  group_by(supplement_consumed, prop_status) %>%
  summarise(Counts = n())

prop.test(prop_tab, 
          alternative = "two.sided",
          correct = FALSE)
```
| | Increased butyrate | Total participants | Percent increase |
|:-----:|:-----:|:-----:|:-----:|
| BRMPS | 54 | 127 | 42.5% |
| LOODAT | 7 | 16 | 43.75% | 

```{r}
# make and save plot (q4_plot)
```

What are your conclusions regarding the reseach question?

Since the p-value for the proportion test was very high, we may suggest that there is no difference in the proportion of propionate reponses between individuals who consumed BRMPS potato (BRMPS) or LOODAT potato supplements.


-----
end